BarnebyLives':' an R package to create herbarium specimen labels and digital data sheets

### 2023-09-29 manuscript/manuscript_APPS.rmd
making labels

```{r load libraries}
library(tidyverse) # data operations
# devtools::install_github('sagesteppe/BarnebyLives')
library(BarnebyLives) # for helping accession collections
```

## Usage
BarnebyLives is run entirely from within Rstudio. 
Data may be read in from any common spreadsheet management system or database connection such as Excel, LibreOffice, OpenOffice, or via the cloud on Googlesheets. 
The latter two options are documented here and in package vignettes, detailed descriptions of the required and suggested input columns are located on the Github page (https://github.com/sagesteppe/BarnebyLives *'Input Data Column Names'*) and over 100 real-world examples are on a Google Sheets accessible from the page.
BarnebyLives is atypical of R packages in that it requires a considerable amount of data to operate (Table 1).
Virtually all of the on-disk memory associated with these data are for storing geo-spatial information, setting up a local instance of the program - at whichever scale a user desires (see Figure XX) is available in the package documentation. Functions which require the on-disk data require a path to the data as an argument. 
Manually supplying the argument allows for the users to judiciously decide a storage location suitable for there needs.  
We anticipate most personal BarnebyLives instances will be less than several gigabytes, and the processing takes relatively little RAM, hence we believe installations can work on hardware as small as Chromebooks, or have the data stored entirely on thumb-drives.
The final steps of Barnebylives, generating the labels require working installations of Rmarkdown, a LaTeX installation (e.g. pdflatex, lualatex, xelatex), and the open source command line tools pdfjam and pdftk.
While these steps are run through bash, we have wrapped them in a R functions which bypass the need to enter the commands to a terminal.
Several commands in BarnebyLives require the output from previous functions, and a workflow which satisfies these requirements is presented in FIGURE XX.

@@

time_geodata_writer <- system.time({ # write out collection as GoogleEarth object
  geodata_writer(data, path = '~/Documents/BL', 
               filename = 'Herbarium_Collections_2023',
               filetype = 'kml')
})

@@

It took roughly three minutes (`r sum( time_trials[time_trials$Group=='Local', 'elapsed'])`s) minutes to run all local steps of BarnebyLives, and a further 16 minutes (`r time_trials[time_trials$Function=='powo_searcher', 'elapsed']`s) to search Plants of the World Online, and `r time_trials[time_trials$Function=='directions_grabber', 'elapsed']` to search Google Maps and write directions to sites.
Nearly all of the local run time is attributable to the spatial operations (spatial: `r sum(tt_local[tt_local$Module=='Geospatial', 'elapsed'])`s, style: `r sum(tt_local[tt_local$Module=='Style', 'elapsed'])`, taxonomic: `r sum(tt_local[tt_local$Module=='Taxonomic', 'elapsed'])`). 
The generation of labels is the most time intensive process and consumed around XX minutes for the rendering, XX to combine individual labels four per single sheet of landscape orientated paper, and XX to combine the XX sheets to a single Portable Document Format (PDF).

@@

```{r import results}
data_p <- readRDS(file = 'data/processed/data_w_POWO_search')

@@

Family_in_positions <- data %>% 
  drop_na(Family)
Family_in <- Family_in_positions %>% 
  filter(Collection_number %in% data_p$Collection_number) %>% 
  arrange(Collection_number) %>% 
  pull(Family)

@@

# NEED TO WORK ON INPUT DATA, INCONSISTENT USE OF AUTHORSHIP IN THE FIRST COLUMN
data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(Full_name != POW_Full_name) %>% 
  select(Collection_number, Full_name, Binomial, POW_Full_name) 
```

>>>>

```{r install package}
devtools::install_github('sagesteppe/BarnebyLives')
```

```{r load libraries}
library(tidyverse) # data operations
library(BarnebyLives) # for helping accession collections
```

## Usage
All steps of BarnebyLives except for label generation are run from within Rstudio. 
Data may be read in from any common spreadsheet management system or database connection such as Excel, LibreOffice, OpenOffice, or via the cloud on Googlesheets. 
The latter two options are documented here and in package vignettes, detailed descriptions of the required and suggested input columns are located on the Github page (https://github.com/sagesteppe/BarnebyLives *'Input Data Column Names'*) and over 100 real-world examples are on a Google Sheets accessible from the page.
BarnebyLives is atypical of R packages in that it requires a considerable amount of data to operate (Table 1).
Virtually all of the on-disk memory associated with these data are for storing geo-spatial information, setting up a local instance of the program - at whichever scale a user desires (see Figure XX) is available in the package documentation. Functions which require the on-disk data require a path to the data as an argument. 
Manually supplying the argument allows for the users to judiciously decide a storage location suitable for there needs.  
We anticipate most personal BarnebyLives instances will be less than several gigabytes, and the processing takes relatively little RAM, hence we believe installations can work on hardware as small as Chromebooks, or have the data stored entirely on thumb-drives.
The final steps of Barnebylives, generating the labels require working installations of Rmarkdown, a LaTeX installation (e.g. pdflatex, lualatex, xelatex), and the open source command line tools pdfjam and pdftk.
While these steps are run through bash, we have wrapped them in a R functions which bypass the need to enter the commands to a terminal.
Several commands in BarnebyLives require the output from previous functions, and a workflow which satisfies these requirements is presented in FIGURE XX.

@@

time_geodata_writer <- system.time({ # write out collection as GoogleEarth object
  geodata_writer(data, path = 'data/processed', 
               filename = 'Herbarium_Collections_2023',
               filetype = 'kml')
})

@@

It took roughly *six* minutes (`r sum( time_trials[time_trials$Group=='Local', 'elapsed'])`s) to run all local steps of BarnebyLives, and roughly 20 minutes (`r time_trials[time_trials$Function=='powo_searcher', 'elapsed']`s) to search Plants of the World Online, and `r time_trials[time_trials$Function=='directions_grabber', 'elapsed']` to search Google Maps and write directions to sites.
Most of the local run time is attributable to the spatial (spatial: `r sum(tt_local[tt_local$Module=='Geospatial', 'elapsed'])`s), and taxonomic operations (`r sum(tt_local[tt_local$Module=='Taxonomic', 'elapsed'])`), style: `r sum(tt_local[tt_local$Module=='Style', 'elapsed'])`. 
The spell check operation of the scientific name accounted for nearly all of the time (`r sum(tt_local[tt_local$Function=='spell_check', 'elapsed'])`) spent performing local taxonomic operations.
The generation of labels consumed around XX minutes for the rendering, XX to combine individual labels four per single sheet of landscape orientated paper, and XX to combine the XX sheets to a single Portable Document Format (PDF).

@@

```{r import results}
data_p <- readRDS(file = 'data/processed/data_w_POWO_search') %>% 
  sf::st_drop_geometry()

@@

Family_in_positions <- data_p %>% 
  drop_na(Family)
Family_in <- Family_in_positions %>% 
  filter(Collection_number %in% data_p$Collection_number) %>% 
  arrange(Collection_number) %>% 
  pull(Family)

@@
